---
title: "Covid91 vaccine study Final2024"
author: "Aubrey Garcia!"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="finalSumr2024.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
```

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

$H_0$:The covid-19 vaccine has no effect on people in different sub groups
$H_A$:The covid-19 vaccine does have an effect on people in different sub groups

## Males

$H_0$:Covid19 is not effective in males

$H_A$:Covid19 is effective in males

### methods
Treatment is a categorical variable

Sex is a categorical variable ..

so we use the cat~cat helper file

### Descriptive results


#### graphical summary
```{r}
males=filter(FinalData,FinalData$sex=="M")
barchartGC(~treatment + infected,data=males)
barchartGC(~infected + treatment,data=males, type="percent")
```


The chart shows in males that those who reccived the placebo had gotten much more positive covid-19 results and those who had gotten the vaccine had a lower amount in positive covid-19 results.



#### numerical summary

```{r}
table1 <- xtabs(~treatment + infected, data=males)
rowPerc(table1)
colPerc(table1)
```


The table shows that 74% of males who received the placebo got Covid-19, and 26% males who got the vaccine got Covid-19

#### Inferential Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```


The inferential results, using a chi-squired test, showed that the p-value is 0.00000000000000022  therefore we reject the null hypothesis because it is less than 0.05 


The Fisher's Exact Test for Count Data shows that the p-vaule is 0.00000000000000022  therefore we reject the null hypothesis because it is less than 0.05. The test also showed that  males have a odd ratio of 2.8 times more likely to get Covid-19 with a placebo compared to those who got the vaccine.  

## Females

$H_0$:Covid19 is not effective in females

$H_A$:Covid19 is effective in females

### methods

Treatment is categorical varible


Sex is a categorical varible

so we use the cat~cat helper file

### Descriptive results

#### graphical summary

```{r}
females=filter(FinalData,FinalData$sex=="F")
barchartGC(~treatment + infected,data=females)
barchartGC(~infected + treatment,data=females, type="percent")
```


The chart shows that in females, those who reccived the placebo had gotten more positive covid-19 results and those who had gotten the vaccine had a much lower amount in positive covid-19 results.


#### numerical summary

```{r}
table1 <- xtabs(~treatment + infected, data=females)
rowPerc(table1)
colPerc(table1)
```

The numerical results showed that 65% of those who got the placebo got covid-19 and that 35% of those who got the vaccine got covid-19.

#### Inferential Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```


The inferential results, using a chi-squired test, showed that the p-value is 0.00000001914  therefore we reject the null hypothesis because it is less than 0.05 

The Fisher's Exact Test for Count Data shows that the p-vaule is 0.00000001246
therefore we reject the null hypothesis because it is less than 0.05. The test also showed that  females have a odd ratio of 1.8 times more likely to get Covid-19 with a placebo compared to those who got the vaccine.  


## LGBTQ

$H_0$:Covid19 is not effective in LGBTQ community

$H_A$:Covid19 is effective in LGBTQ community


### methods

Treatment is a categorical variable

LGBTQ is a categorical variable

So we use the cat~cat helper file

### Descriptive results

#### graphical summary

```{r}
LGBTQ=filter(FinalData,FinalData$LGBTQ=="gay")
barchartGC(~treatment + infected,data=LGBTQ)
barchartGC(~infected + treatment,data=LGBTQ, type="percent")
```

The table shows that in the LGBTQ community there was more positive covid-19 results among those who has received the placebo and the ones that had gotten the vaccine and much lower positive covid-19 results.

#### numerical summary

```{r}
table1 <- xtabs(~treatment + infected, data=LGBTQ)
rowPerc(table1)
colPerc(table1)
```

The numerical results showed that 80% of those who got the placebo got covid-19 and that 20% of those who got the vaccine got covid-19.


#### Inferential Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The inferential results, using a chi-squired test, showed that the p-value is 0.00000000000000022  therefore we reject the null hypothesis because it is less than 0.05 

The Fisher's Exact Test for Count Data shows that the p-vaule is 0.00000000000000022 therefore we reject the null hypothesis because it is less than 0.05. The test also showed that lgbtq have a odd ratio of 4.2 times more likely to get Covid-19 with a placebo compared to those who got the vaccine. 


## Druggies

$H_0$:Covid19 is not effective in drug users.

$H_A$:Covid19 is effective in drug users.

### methods

Treatment is a categorical variable

Drug users is a categorical variable 

so we use the cat~cat helper file

### Descriptive results

#### graphical summary

```{r}
DrugUser=filter(FinalData)
barchartGC(~treatment + infected,data=DrugUser)
barchartGC(~infected + treatment,data=DrugUser, type="percent")
```

The table shows that among the DrugUsers there was more positive covid-19 results among those who has received the placebo and the ones that had gotten the vaccine and much lower positive covid-19 results.


#### numerical summary

```{r}
table1 <- xtabs(~treatment + infected, data=DrugUser)
rowPerc(table1)
colPerc(table1)
```

The numerical results showed that 69% of those who got the placebo got covid-19 and that 31% of those who got the vaccine got covid-19.


#### Inferential Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The inferential results, using a chi-squired test, showed that the p-value is 0.00000000000000022  therefore we reject the null hypothesis because it is less than 0.05 

The Fisher's Exact Test for Count Data shows that the p-vaule is 0.00000000000000022 therefore we reject the null hypothesis because it is less than 0.05. The test also showed that drugusers have a odd ratio of 2.2 times more likely to get Covid-19 with a placebo compared to those who got the vaccine. 

# Overall Results and Conclusions